---
title: "STA 135 - Homework 2"
author: "Gianni Spiga"
date: "4/20/2022"
output:
  pdf_document: default
  html_document: default
---

# Problem 3

a.)
```{r}
scores <- read.table("~/GitHub/STA135/Homework/HW2/T5-2.dat")

meanalt <- c(500, 50, 30)

library(DescTools)
HotellingsT2Test(scores, mu = meanalt, test = 'chi')

#Finding the critical point

n = nrow(scores)
p = ncol(scores)

Fstat = qf(0.05, p, n - p, lower.tail =  FALSE)
critval <- ((n -1) * p) / (n - p) * Fstat

```
Given that our $T^2$ value is much larger than our critical value as well as our p-value is extremely close to zero, we reject $H_0$. We have reason to believe that students are scoring differently than the reported average scores. 

b.)
```{r}
S <- cov(scores)
eigVal <- eigen(S)$values
eigVec <- eigen(S)$vectors



cat("The directions of the corresponding axis are \n")
eigVec
cat("\n")
cat("The lengths of the axis are \n")
sqrt(eigVal) * sqrt((1/n) * critval)

```
c.)
```{r}
qqnorm(scores[,1]); qqline(scores[,1])
qqnorm(scores[,2]); qqline(scores[,2])
qqnorm(scores[,3]); qqline(scores[,3]) 

library(ggplot2)
library(ggExtra)
library(GGally)

#Pairwise plots
ggpairs(scores)

```
\newline

# Problem 4
P 5.21 
```{r}
bone <- read.table("~/GitHub/STA135/Homework/HW2/T1-8.dat")
bone

#Bonferroni
tstat <- qt(0.05/(2*ncol(bone)), nrow(bone) - 1) 
tstat

upperBon <- colMeans(bone) - tstat * sqrt(diag(var(bone))/nrow(bone))
lowerBon <- colMeans(bone) + tstat * sqrt(diag(var(bone))/nrow(bone))

bonf <- matrix(c(lowerBon, upperBon), nrow= 6, ncol = 2)
bonf

n <- nrow(bone)
p <- ncol(bone)

# T^2 Interval
T2stat <- sqrt((p * (n- 1))/(n- p) * qf(0.05, p, n - p, lower.tail = FALSE))

lowerT2 <- colMeans(bone) - T2stat * sqrt(diag(var(bone))/nrow(bone))
upperT2 <- colMeans(bone) + T2stat * sqrt(diag(var(bone))/nrow(bone))

T2int <- matrix(c(lowerT2, upperT2), nrow= 6, ncol = 2)
T2int
```
We can see in this example, that the intervals for our simultaneous Bonferroni intervals are more narrow than the simulataneous $T^2$ intervals. 
