---
title: "Final Project"
author: "Gianni Spiga"
date: "6/6/2022"
output: html_document
---

## P1
a.)
```{r}
#Picking a very random seed 
set.seed(503650269)

skull <- read.table("~/GitHub/STA135/Final Project/Data/T6-13.dat")
colnames(skull) <- c("MaxBreath", "BasHeight", "BasLength", "NasHeight", "TimePeriod")
skull


skullSamp1 <- skull[sample(1:30, 15), ]
skullSamp1

skullSamp2 <- skull[sample(31:60, 15), ]
skullSamp2

skullSamp3 <- skull[sample(61:90, 15), ]
skullSamp3
```
```{r}
skullSamp <- rbind(skullSamp1, skullSamp2, skullSamp3)
#skullSamp <- skullSamp[1:(length(skullSamp) - 1)]
skullSamp

model <- manova(cbind(skullSamp$MaxBreath, skullSamp$BasHeight, skullSamp$BasLength, skullSamp$NasHeight) ~ skullSamp$TimePeriod, data = skullSamp)
model
summary(model)
results <- HoRM::SSCP.fn(model)
results
```

```{r}
# Sum of all residuals 
#SSres  <- sum(model$residuals[,1]^2) + sum(model$residuals[,2]^2) + sum(model$residuals[,3]^2) + sum(model$residuals[,4]^2)
SStrt <- results$SSCPR
SSres <- results$SSCPE
SStot <- results$SSCPTO

wilks <- det(SSres) / det(SStot)
wilks


#Test stat
testStat <- qf(0.05, df1 = 2 * ncol(skullSamp) - 1, df2 = 2*(nrow(skullSamp) - 4 - 2), lower.tail = FALSE)
testStat

#Last row of table on pg 303, g = 3, p >_ 1
critval <- ((45 - 4 - 2) / 4) * ((1 - sqrt(wilks)) / sqrt(wilks)) 
critval

testStat > critval
```

Since our test statistic is larger than critical value, we reject $H_0$, concluding that at least one time period had significantly different skull dimensions than the others. 

b.)

```{r}
colMeans(skullSamp1)
colMeans(skullSamp2)
colMeans(skullSamp3)
```
We can see from the column means that the average Base height stays about the same, however in other variavles such as Max Breath and Nas Height, time period 2 is less than time periods 1 and 3. 

## P2
a.)
```{r}
library(readxl)
word <-
  read_excel("~/GitHub/STA135/Final Project/Data/wordparity.xlsx")
#Randomly sample 20 columns
word <- word[sample(nrow(word), 20),]
word


#Hotellings for same v different
# Different
diff <- word[,c(1,3)]
xbarDiff <- colMeans(diff)
S1diff <- cov(diff)

# Same
same <- word[,c(2,4)]
xbarSame <- colMeans(same)
S2same <- cov(same)

Spool <- ((nrow(diff) - 1) / (nrow(diff) + nrow(same) - 2)) * S1diff  + ((nrow(same) - 1) / (nrow(diff) + nrow(same) - 2)) * S2same

T2words <- (xbarDiff - xbarSame) %*% solve(((1 / nrow(diff)) +  (1/nrow(same))) * Spool) %*% (xbarDiff - xbarSame)
T2words

#Now calculate crit val
# We will assume alpha = 0.05
critval_words <- ((nrow(diff) + nrow(same) - 2) * ncol(diff)) / (nrow(diff) + nrow(same) - ncol(word) - 1) * qf(0.05, ncol(diff), (nrow(diff) + nrow(same) - ncol(word) - 1), lower.tail = FALSE)
critval_words
```
Since our $T^2$ is greater than our critical value, we  reject our null hypothesis $H_0$. We have evidence to believe that there is a significant difference between the means of reaction times for type 'same' and type 'different'. 

b.)
```{r}
#Hotellings for Arabic v word
# Different
written <- word[,c(1, 2)]
xbarWrit <- colMeans(written)
S1Writ <- cov(written)

# Same
arab <- word[,c(3,4)]
xbarArab <- colMeans(arab)
S2Arab <- cov(arab)

Spool <- ((nrow(written) - 1) / (nrow(written) + nrow(arab) - 2)) * S1Writ  + ((nrow(arab) - 1) / (nrow(written) + nrow(arab) - 2)) * S2Arab

T2type <- (xbarWrit - xbarArab) %*% solve(((1 / nrow(written)) +  (1/nrow(arab))) * Spool) %*% (xbarWrit - xbarArab)
T2type

#Now calculate crit val
# We will assume alpha = 0.05
critval_type <- ((nrow(written) + nrow(arab) - 2) * ncol(written)) / (nrow(written) + nrow(arab) - ncol(written) - 1) * qf(0.05, ncol(written), (nrow(written) + nrow(arab) - ncol(written) - 1), lower.tail = FALSE)
critval_words

```
## P3
```{r}
census <- read.table("~/GitHub/STA135/Final Project/Data/T8-5.dat")
census


```

