---
title: "Homework 4"
author: "Gianni Spiga"
date: "5/10/2022"
output:
  pdf_document: default
  html_document: default
---

### P1 
6.8.)
a.)
```{r}
tr1 <- matrix(c(6, 7, 5, 9, 8, 6, 4, 9, 7, 9), nrow = 2, ncol = 5)
tr1

tr2 <- matrix(c(3, 3, 1, 6, 2, 3), nrow = 2, ncol = 3)
tr2

tr3 <- matrix(c(2, 3, 5, 1, 3, 1, 2, 3), nrow = 2, ncol = 4)
tr3

xbar1 <- rowMeans(tr1)
xbar1

xbar2 <- rowMeans(tr2)
xbar2

xbar3 <- rowMeans(tr3)
xbar3


#xl1 <- rbind(tr1[1,], tr2[2,], tr3[3,])

xbar <-
  c((ncol(tr1) * xbar1[1] + ncol(tr2) * xbar2[1] + ncol(tr3) * xbar3[1]) / (ncol(tr1) + ncol(tr2) + ncol(tr3)),
    ((ncol(tr1) * xbar1[2]) + (ncol(tr2) * xbar2[2]) + (ncol(tr3) * xbar3[2])) / (ncol(tr1) + ncol(tr2) + ncol(tr3))
  )
xbar


#cat("=", )

#xbarl1 <- matrix(4, nrow = 3, ncol = 5) +
# cat("=") 
# matrix(4, nrow = 3, ncol = 5)
# print('+' )
# matrix(c(2,-2,-1), nrow = 3, ncol =5)
# print('+')
# matrix(c(2,-2,-1), nrow = 3, ncol =5)
```


$$
x_{1j} = \begin{pmatrix}
4 & 4 & 4 & 4 & 4 \\
4 & 4 & 4 &\\
4 &4 &4 &4 &\\
\end{pmatrix} + \begin{pmatrix}
2 & 2 & 2 & 2 & 2 \\
-2 & -2 & -2 &\\
-1 & -1 & -1 & -1 &\\
\end{pmatrix} + \begin{pmatrix}
0 & -1 & 2 & -2 & 1 \\
1 & -1 & 0 &\\
-1 & 2 & 0 & -1 &\\
\end{pmatrix}
$$
$$
x_{2j} = \begin{pmatrix}
5 & 5 & 5 & 5 & 5 \\
5 & 5 & 5 &\\
5 &5 &5 &5 &\\
\end{pmatrix} + \begin{pmatrix}
3 & 3 & 3 & 3 & 3 \\
-1 & -1 & -1 &\\
-3 & -3 & -3 & -3 &\\
\end{pmatrix} + \begin{pmatrix}
-1 & 1 & -2 & 1 & 1 \\
-1 & 2 & -1 &\\
-1 & 2 & 0 & -1 &\\
\end{pmatrix}
$$

b.)
Variable 1
```{r}
mat1 <- rbind(tr1[1,], c(tr2[1,], rep(0, 2)), c(tr3[1,], 0))
mat1

SSobs1 <- sum(mat1 ^ 2)
SSobs1

SSmean1 <- xbar[1] ^ 2 * 12
SSmean1

SStrt1 <-
  ncol(tr1) * (mean(tr1[1,] - xbar[1])) ^ 2 + ncol(tr2) * (mean(tr2[1,] - xbar[1])) ^
  2 + ncol(tr3) * (mean(tr3[1,] - xbar[1])) ^ 2
SStrt1

SSres1 <-
  sum((mat1 - rbind(rep(xbar1[1], 5),
                    c(rep(
                      xbar2[1], 3
                    ), 0 , 0),
                    c(rep(
                      xbar3[1], 4
                    ), 0))) ^ 2)
SSres1

SStotal1 <- SSobs1 - SSmean1
SStotal1
```
Variable 2 
```{r}
mat2 <- rbind(tr1[2,], c(tr2[2,], rep(0, 2)), c(tr3[2,], 0))
SSobs2 <- sum(mat2 ^ 2)
SSobs2

SSmean2 <- xbar[2] ^ 2 * 12
SSmean2

SStrt2 <-
  ncol(tr1) * (mean(tr1[2,] - xbar[2])) ^ 2 + ncol(tr2) * (mean(tr2[2,] - xbar[2])) ^
  2 + ncol(tr3) * (mean(tr3[2,] - xbar[2])) ^ 2
SStrt2

SSres2 <-
  sum((mat2 - rbind(rep(xbar1[2], 5),
                    c(rep(
                      xbar2[2], 3
                    ), 0 , 0),
                    c(rep(
                      xbar3[2], 4
                    ), 0))) ^ 2)
SSres2

SStotal2 <- SSobs2 - SSmean2
SStotal2
```

Cross Product 
```{r}
SSobs_x <- sum(mat1 * mat2)
SSobs_x

SSmean_x <- xbar[1] * xbar[2] * 12
SSmean_x

SStrt_x <- ncol(tr1) *  (mean(tr1[1,] - xbar[1]))* (mean(tr1[2,] - xbar[2])) + ncol(tr2) * (mean(tr2[1,] - xbar[1])) * (mean(tr2[2,] - xbar[2])) + ncol(tr3) * (mean(tr3[1,] - xbar[1]))* (mean(tr3[2,] - xbar[2]))
SStrt_x

SSres_x <- sum((mat2 - rbind(rep(xbar1[2], 5), c(rep( xbar2[2], 3), 0 , 0),
                    c(rep(xbar3[2], 4
                    ), 0))) * (mat1 - rbind(rep(xbar1[1], 5),
                    c(rep(
                      xbar2[1], 3
                    ), 0 , 0),
                    c(rep(
                      xbar3[1], 4
                    ), 0))))
SSres_x

SStotal_x <- SSobs_x - SSmean_x
SStotal_x
```

```{r}
cat("Treatment \n")
B <- matrix(c(SStrt1, SStrt_x, SStrt_x, SStrt2), nrow = 2, ncol = 2)
B
dfB <- nrow(mat1)-1
dfB

cat("Residuals \n")
W <- matrix(c(SSres1, SSres_x, SSres_x, SSres2), nrow = 2, ncol = 2)
W
dfW <- 12 - 3
dfW

cat("Total \n")
B + W
#df
12 - 1

```

c.)
```{r}
WLambda <- det(W) / det(W + B)
WLambda

#Test stat
((12 - nrow(mat1) - 1) / (nrow(mat1) - 1)) * ((1 - sqrt(WLambda)) / sqrt(WLambda))

#Critical value
qf(0.01, df1 = 2 * nrow(tr1), df2 = 2 *(12 - nrow(tr1) - 2), lower.tail = FALSE)
```

Since our critical value is less than our test statistic, we reject $H_0$ when $\alpha = 0.01$. We have statistical evidence to conclude at least one $\tau_i$ is not 0. 

Barlett's correction
```{r}
#Test statistic
- (12 - 1 - ((nrow(tr1) + nrow(mat1))/2)) * log(WLambda)

#Critical Value
qchisq(0.01, df = nrow(tr1) * (nrow(mat1)-1), lower.tail = FALSE)

```
Since our critical value is once again less than our test statistic, we reject $H_0$. This conclusion lines up with the conclusion we made using Wilk's Lambda. 
While both are adequite in the case of small sample sizes, Bartlett's correction improves accuracy for large sample sizes. 

P2
6.13.)
a.) and b.)

```{r}
# tr1 <- matrix(c(6, 4, 8, 2, 3, -3, 4, -4, -3, -4, 3, -4), nrow = 3, ncol = 4, byrow = TRUE)
# tr1
# 
# tr2 <- matrix(c(8,6,12,6,8,2,3,3,2,-5,-3,-6), nrow = 3, ncol = 4, byrow = TRUE)
# tr2
# 
# xbar1 <- mean(tr1)
# xbar1
# 
# xbar2 <- mean(tr2)
# xbar2
# 
# xbar <- matrix(xbar1, nrow = 3, ncol = 4)
# xbar
# 
# factef1 <- rowMeans(tr1) - matrix(xbar1, nrow = 3, ncol = 4)
# factef1
# 
# factef2 <- matrix(c(rep(colMeans(tr1) - xbar1, 3)), nrow = 3, ncol = 4, byrow = TRUE) 
# factef2
# 
# resid1 <- tr1 - factef1 - factef2 - xbar
# resid1

```

```{r}
tr1 <- matrix(c(6, 4, 8, 2, 3, -3, 4, -4, -3, -4, 3, -4), nrow = 3, ncol = 4, byrow = TRUE)
tr1

tr2 <- matrix(c(8,6,12,6,8,2,3,3,2,-5,-3,-6), nrow = 3, ncol = 4, byrow = TRUE)
tr2

xbar1 <- mean(tr1)
xbar1

xbar2 <- mean(tr2)
xbar2

xbar <- matrix(xbar1, nrow = 3, ncol = 4)
xbar

factef1_1 <- rowMeans(tr1) - matrix(xbar1, nrow = 3, ncol = 4)
factef1_1

factef2_1 <- matrix(c(rep(colMeans(tr1) - xbar1, 3)), nrow = 3, ncol = 4, byrow = TRUE) 
factef2_1

resid1 <- tr1 - factef1_1 - factef2_1 - xbar
resid1

SSmean1 <- sum(xbar^2)
SSmean1

SSfac1_1 <- sum(factef1_1^2)
SSfac1_1

SSfac2_1 <- sum(factef2_1^2)
SSfac2_1

SSres1 <- sum(resid1 ^2)
SSres1

SStol <- SSmean1 + SSfac1_1 + SSfac2_1 + SSres1
SStol
```
```{r}
xbar <- matrix(xbar2, nrow = 3, ncol = 4)
xbar

factef1 <- rowMeans(tr2) - matrix(xbar2, nrow = 3, ncol = 4)
factef1

factef2 <- matrix(c(rep(colMeans(tr2) - xbar2, 3)), nrow = 3, ncol = 4, byrow = TRUE) 
factef2

resid2 <- tr2 - factef1 - factef2 - xbar
resid2


SSmean2 <- sum(xbar^2)
SSmean2

SSfac1_2 <- sum(factef1^2)
SSfac1_2

SSfac2_2 <- sum(factef2^2)
SSfac2_2

SSres2 <- sum(resid2 ^2)
SSres2

SStol <- SSmean2 + SSfac1_2 + SSfac2_2 + SSres2
SStol

```
```{r}
SCPmean <-
  sum(matrix(xbar2, nrow = 3, ncol = 4) * matrix(xbar1, nrow = 3, ncol = 4))
SCPmean

SCPfac1 <-
  sum((rowMeans(tr2) - matrix(xbar2, nrow = 3, ncol = 4)) * (rowMeans(tr1) - matrix(xbar1, nrow = 3, ncol = 4)))
SCPfac1

SCPfac2 <- 
  sum((matrix(c(rep(colMeans(tr1) - xbar1, 3)), nrow = 3, ncol = 4, byrow = TRUE) * (matrix(c(rep(colMeans(tr2) - xbar2, 3)), nrow = 3, ncol = 4, byrow = TRUE))))
SCPfac2

SCPres <- sum(resid1 * resid2)
SCPres

SCPtot <- SCPmean + SCPfac1 + SCPfac2 + SCPres
SCPtot
```
c.)
```{r}
# Manova table
cat("Factor 1 \n")
fac1 <- matrix(c(SSfac1_1, SCPfac1, SCPfac1, SSfac1_2), nrow = 2, ncol = 2)
fac1
dfFac1 <- nrow(tr1)-1
dfFac1

cat("Factor 2 \n")
fac2 <- matrix(c(SSfac2_1, SCPfac2, SCPfac2, SSfac2_2), nrow = 2, ncol = 2)
fac2
dfFac2 <- ncol(tr1)-1
dfFac2

cat("Residuals \n")
resid <- matrix(c(SSres1, SCPres, SCPres, SSres2), nrow = 2, ncol = 2)
resid
dfresid <- (nrow(tr1)-1) * (ncol(tr1)-1)
dfresid

cat("Total \n")
fac1 + fac2 + resid
#df
nrow(tr1) * ncol(tr1) - 1
```
d.)
```{r}
-1 * ((nrow(tr1) - 1)*(ncol(tr1) - 1) - (((2 + 1) - (nrow(tr1) - 1))/2) ) * log(det(resid)/ det(fac1 + resid))

#Critical Value
qchisq(.05, 4, lower.tail = FALSE)

```
Since our test statistic is greater than our critical value, we reject $H_0$, supporting the claim that there are factor 1 effects. 

```{r}
-1 * ((nrow(tr2) - 1)*(ncol(tr2) - 1) - (((2 + 1) - (ncol(tr2) - 1))/2) ) * log(det(resid)/ det(fac2 + resid))

#Critical Value
qchisq(.05, (nrow(tr2) - 1)*(ncol(tr2) - 1), lower.tail = FALSE)

```
We once again reject the null hypothesis and conclude there are factor 2 effects as well. 

6.14.)
```{r}
tr1 <- matrix(c(14, 6, 8, 16, 1, 5, 0, 2, 3, -2, -11, -6), nrow = 3, ncol = 4, byrow = TRUE)
tr1

tr2 <- matrix(c(8, 2, 2, -4, 6, 12, 15, 7 , -2, 7, 1, 6), nrow = 3, ncol = 4, byrow = TRUE)
tr2

xbar1 <- mean(tr1)
xbar1

xbar2 <- mean(tr2)
xbar2

xbar <- matrix(xbar1, nrow = 3, ncol = 4)
xbar

factef1_1 <- rowMeans(tr1) - matrix(xbar1, nrow = 3, ncol = 4)
factef1_1

factef2_1 <- matrix(c(rep(colMeans(tr1) - xbar1, 3)), nrow = 3, ncol = 4, byrow = TRUE) 
factef2_1

resid1 <- tr1 - factef1_1 - factef2_1 - xbar
resid1

SSmean1 <- sum(xbar^2)
SSmean1

SSfac1_1 <- sum(factef1_1^2)
SSfac1_1

SSfac2_1 <- sum(factef2_1^2)
SSfac2_1

SSres1 <- sum(resid1 ^2)
SSres1

SStol <- SSmean1 + SSfac1_1 + SSfac2_1 + SSres1
SStol
```

```{r, echo = FALSE}
# xbar <- matrix(xbar2, nrow = 3, ncol = 4)
# xbar
# 
# factef1 <- rowMeans(tr2) - matrix(xbar2, nrow = 3, ncol = 4)
# factef1
# 
# factef2 <- matrix(c(rep(colMeans(tr2) - xbar2, 3)), nrow = 3, ncol = 4, byrow = TRUE) 
# factef2
# 
# resid2 <- tr2 - factef1 - factef2 - xbar
# resid2
# 
# 
# SSmean2 <- sum(xbar^2)
# SSmean2
# 
# SSfac1_2 <- sum(factef1^2)
# SSfac1_2
# 
# SSfac2_2 <- sum(factef2^2)
# SSfac2_2
# 
# SSres2 <- sum(resid2 ^2)
# SSres2
# 
# SStol <- SSmean2 + SSfac1_2 + SSfac2_2 + SSres2
# SStol
```
```{r}
SCPmean <-
  sum(matrix(xbar2, nrow = 3, ncol = 4) * matrix(xbar1, nrow = 3, ncol = 4))
SCPmean

SCPfac1 <-
  sum((rowMeans(tr2) - matrix(xbar2, nrow = 3, ncol = 4)) * (rowMeans(tr1) - matrix(xbar1, nrow = 3, ncol = 4)))
SCPfac1

SCPfac2 <- 
  sum((matrix(c(rep(colMeans(tr1) - xbar1, 3)), nrow = 3, ncol = 4, byrow = TRUE) * (matrix(c(rep(colMeans(tr2) - xbar2, 3)), nrow = 3, ncol = 4, byrow = TRUE))))
SCPfac2

SCPres <- sum(resid1 * resid2)
SCPres

SCPtot <- SCPmean + SCPfac1 + SCPfac2 + SCPres
SCPtot
```

b.)
```{r}
# tr1_13 <- matrix(c(6, 4, 8, 2, 3, -3, 4, -4, -3, -4, 3, -4), nrow = 3, ncol = 4, byrow = TRUE)
# tr1_13
# 
# tr2_13 <- matrix(c(8,6,12,6,8,2,3,3,2,-5,-3,-6), nrow = 3, ncol = 4, byrow = TRUE)
# tr2_13
# 
# tr1_14 <- matrix(c(14, 6, 8, 16, 1, 5, 0, 2, 3, -2, -11, -6), nrow = 3, ncol = 4, byrow = TRUE)
# tr1_14
# 
# tr2_14 <- matrix(c(8, 2, 2, -4, 6, 12, 15, 7 , -2, 7, 1, 6), nrow = 3, ncol = 4, byrow = TRUE)
# tr2_14
# 
# 
# tr1 <- (tr1_13 + tr1_14) #/ 2
# tr1
# 
# tr2 <- (tr2_13 + tr2_14) #/ 2
# tr2
# 
# 
# 
# tr1 <- matrix(c(14, 6, 8, 16, 1, 5, 0, 2, 3, -2, -11, -6), nrow = 3, ncol = 4, byrow = TRUE)
# tr1
# tr2 <- matrix(c(8, 2, 2, -4, 6, 12, 15, 7 , -2, 7, 1, 6), nrow = 3, ncol = 4, byrow = TRUE)
# tr2
```
```{r}
xbar1 <- mean(tr1)
xbar1

xbar2 <- mean(tr2)
xbar2

xbar <- matrix(xbar1, nrow = 3, ncol = 4)
xbar

factef1_1 <- rowMeans(tr1) - matrix(xbar1, nrow = 3, ncol = 4)
factef1_1

factef2_1 <- matrix(c(rep(colMeans(tr1) - xbar1, 3)), nrow = 3, ncol = 4, byrow = TRUE) 
factef2_1

resid1 <- tr1 - factef1_1 - factef2_1 - xbar
resid1

SSmean1 <- sum(xbar^2)
SSmean1

SSfac1_1 <- sum(factef1_1^2)
SSfac1_1

SSfac2_1 <- sum(factef2_1^2)
SSfac2_1

SSres1 <- sum(resid1 ^2)
SSres1

SStol <- SSmean1 + SSfac1_1 + SSfac2_1 + SSres1
SStol
```

```{r}
xbar <- matrix(xbar2, nrow = 3, ncol = 4)
xbar

factef1 <- rowMeans(tr2) - matrix(xbar2, nrow = 3, ncol = 4)
factef1

factef2 <- matrix(c(rep(colMeans(tr2) - xbar2, 3)), nrow = 3, ncol = 4, byrow = TRUE) 
factef2

resid2 <- tr2 - factef1 - factef2 - xbar
resid2


SSmean2 <- sum(xbar^2)
SSmean2

SSfac1_2 <- sum(factef1^2)
SSfac1_2

SSfac2_2 <- sum(factef2^2)
SSfac2_2

SSres2 <- sum(resid2 ^2)
SSres2

SStol <- SSmean2 + SSfac1_2 + SSfac2_2 + SSres2
SStol
```
```{r}
xbar <- matrix(xbar2, nrow = 3, ncol = 4)
xbar

factef1 <- rowMeans(tr2) - matrix(xbar2, nrow = 3, ncol = 4)
factef1

factef2 <- matrix(c(rep(colMeans(tr2) - xbar2, 3)), nrow = 3, ncol = 4, byrow = TRUE) 
factef2

resid2 <- tr2 - factef1 - factef2 - xbar
resid2


SSmean2 <- sum(xbar^2)
SSmean2

SSfac1_2 <- sum(factef1^2)
SSfac1_2

SSfac2_2 <- sum(factef2^2)
SSfac2_2

SSres2 <- sum(resid2 ^2)
SSres2

SStol <- SSmean2 + SSfac1_2 + SSfac2_2 + SSres2
SStol

```
c.)
```{r}
-1 * ((nrow(tr1) - 1)*(ncol(tr1) - 1) - (((2 + 1) - (nrow(tr1) - 1))/2) ) * log(det(resid)/ det(fac1 + resid))

#Critical Value
qchisq(.05, 4, lower.tail = FALSE)


-1 * ((nrow(tr2) - 1)*(ncol(tr2) - 1) - (((2 + 1) - (ncol(tr2) - 1))/2) ) * log(det(resid)/ det(fac2 + resid))

#Critical Value
qchisq(.05, (nrow(tr2) - 1)*(ncol(tr2) - 1), lower.tail = FALSE)

```
We reject the null hypothesis for factor 1  however we fail to reject factor 2. We can conclude that the factor 1 effect is non zero but we cannot conclude that factor 2 effects are non zero. 

```{r}
cat("Factor 1 \n")
fac1 <- matrix(c(SSfac1_1, SCPfac1, SCPfac1, SSfac1_2), nrow = 2, ncol = 2)
fac1
dfFac1 <- nrow(tr1)-1
dfFac1

cat("Factor 2 \n")
fac2 <- matrix(c(SSfac2_1, SCPfac2, SCPfac2, SSfac2_2), nrow = 2, ncol = 2)
fac2
dfFac2 <- ncol(tr1)-1
dfFac2

cat("Residuals \n")
resid <- matrix(c(SSres1, SCPres, SCPres, SSres2), nrow = 2, ncol = 2)
resid
dfresid <- (nrow(tr1)-1) * (ncol(tr1)-1)
dfresid

cat("Total \n")
fac1 + fac2 + resid
#df
nrow(tr1) * ncol(tr1) - 1
```
# 6.19.)

a.)
```{r}
X <- read.csv("T6-10.DAT", header = FALSE, sep = "")
names(X) <- c("fuel", "repair", "capital", "truck")

X1 <- X[X$truck == "gasoline", 1:3]
X2 <- X[X$truck == "diesel", 1:3]

n1 <- nrow(X1)
n2 <- nrow(X2)
p <- ncol(X1)

null <- rep(0, p)
xbar1 <- colMeans(X1)
xbar2 <- colMeans(X2)

S1 <- cov(X1)
S2 <- cov(X2)
Spool <- ((n1-1)*S1+(n2-1)*S2)/(n1+n2-2)

T2 <- (n1*n2/(n1+n2))*t(xbar1-xbar2-null)%*%solve(Spool)%*%(xbar1-xbar2-null)
critval <- (n1 + n2 - 2)*p/(n1 + n2 - p - 1)*qf(0.01, p, n1 + n2 - p - 1, lower.tail = FALSE)

cat("Hotellings T2 is:", T2)
cat("\nCritical value is:", critval)

```

Since our critical value is less than our $T^2$ value, we reject $H_0$, concluding there is statistical evidence that the cost vectors between groups are not zero vectors. 

b.)
```{r}
solve(Spool)%*%(xbar1 - xbar2)
```
We can see from our solution vector that the variable with the highest absolute influence would be capital, followed by fuel, and finally repair. 

c.)
```{r}
lowerCI <- t(diag(p))%*%(xbar1 - xbar2) - sqrt(critval) * sqrt(diag((t(diag(p))%*%Spool%*%diag(p))*(1/n1 + 1/n2)))
upperCI <- t(diag(p))%*%(xbar1 - xbar2) + sqrt(critval) * sqrt(diag((t(diag(p))%*%Spool%*%diag(p))*(1/n1 + 1/n2)))

results <- as.data.frame(cbind(lowerCI, upperCI), 
                     row.names = c("Fuel", "Repair", "Capital"))
results
```

d.)
```{r}
Xred <- X[-c(9, 21),]

X1red <- Xred[Xred$truck=="gasoline",1:3]
X2red <- Xred[Xred$truck=="diesel",1:3]

n1 <- nrow(X1red)
n2 <- nrow(X2red)

xbar1 <- colMeans(X1red)
xbar2 <- colMeans(X2red)

S1 <- cov(X1red)
S2 <- cov(X2red)
Spool <- ((n1-1)*S1+(n2-1)*S2)/(n1+n2-2)

T2 <- (n1*n2/(n1+n2))*t(xbar1-xbar2-null)%*%solve(Spool)%*%(xbar1-xbar2-null)
critval <- (n1 + n2 - 2)*p/(n1 + n2 - p - 1)*qf(0.01, p, n1 + n2 - p - 1, lower.tail = FALSE)

cat("Hotellings T2 is:", T2)
cat("\nCritical value is:", critval)
```
We can see here that our critical value is less than our $T^2$, we we still reject the null hypothesis, that is that the mean difference in cost vectors is not the zero vector. 
